Water Management

### 2020-02-11 Paper/paper.tex
working on section bellman equation and implementation

@@

@@

In a first step, we want to compute the value function $V$ for each state $s$. In order to do so, we need to calculate an auxiliary matrix $aux$ that contains the combined utilities of the farmers and the recreational users. Obviously, the $aux$ matrix can only a finite, discrete amount of values. In this work, a resolution of $n = 1000$ different water levels and irrigation amounts should suffice. Thus we discretize both water level $s$ and irrigation amount $x$ as vectors
$$
s = \left( \begin{array}{c}0\\...\\7\end{array}\right), \hspace{3mm}x = \left( \begin{array}{c}0\\...\\7\end{array}\right)
$$
Note that the maximum reservoir capacity is $7$ and $|s|=|x| = n = 1000$.

For a combined utility of the farmers and the recreational users  $u(s_i,x_j) = F(x_j) + G(s_i,x_j)$, depending on the water level $s_i$ and the amount of irrigation $x_i$ we get an auxiliary matrix
$$
aux = \left[ \begin{array}{ccccc}                                u(s_1, x_1)& \varnothing & \varnothing & ... & \varnothing \\                                               u(s_2, x_1)& u(s_2, x_2)& \varnothing & ... & \varnothing \\                                               u(s_3, x_1)& ... & u(s_3, x_3)& ... & \varnothing \\                                              ... & ... & ... & ... & \varnothing \\                                              u(s_n, x_1)& ... & ... & ... & u(s_n, x_n)\\                                               \end{array}\right]\\\\\\+\beta*\left[ \begin{array}{ccccc}                                V'(s_1') & \varnothing & \varnothing & ... & \varnothing \\                                               V'(s_1') & V'(s_2') & \varnothing & ... & \varnothing \\                                               V'(s_1') & ... & V'(s_3') & ... & \varnothing \\                                              ... & ... & ... & ... & \varnothing \\                                              V'(s_1') & ... & ... & ... & V'(s_n') \\                                               \end{array}\right]
$$
where upper right triangle of this matrix will be empty, as the amount of water, that can be taken from the reservoir must be smaller or equal to the amount of water in the reservoir.

@@

$$V(s) = \underset{x}{max}(u(s,x) + \beta*V'(s')))$$

which equals to finding the column $aux$ that maximizes the expected value over all states (TODO: kann man das so sagen?). We repeat this step over and over again until the our value function $V$ converges.

>>>>

\usepackage{breqn} % for line breaks in equations

@@

\subsection{The Bellman Equation}

The so-called "Bellman Equation" answers the question, assuming a given state, what long-term utility can be expected by taking the best possible action now and at each subsequent step - or simply put: "What is the value $V$ of a state $s$?".

Mathematically, we can express this as  
\begin{equation}
V(s) = \max_{x}(u(s,x) + \beta*V'(s')))
\end{equation}

where $u(s,x)$ is the utility to gain from the control variable (action) $x$ in state $s$ [TODO: QUOTE Seminar Week 3].

So the value of the control variable to be set is the one that maximizes both the utility of the current state plus (recursively) the value $V$ of the next state $s'$ which is discounted by a factor of $\beta$.

@@

In a first step, we want to compute the value function $V$ for each state $s$. In order to do so, we need to calculate an auxiliary matrix $aux$ that contains the combined utilities of the farmers and the recreational users. Obviously, the $aux$ matrix can only a finite, discrete amount of values. In this work, a resolution of $n = 1000$ different water levels and irrigation amounts should suffice. Thus we discretize both water level $s$ and irrigation amount $x$ as vectors

	\begin{equation}
s = \left( \begin{array}{c}0\\...\\7\end{array}\right),\hspace{2mm}x = \left( \begin{array}{c}0\\...\\7\end{array}\right)
\end{equation}

Note that the maximum reservoir capacity is $7$ and $|s|=|x| = n = 1000$.

For a combined utility of the farmers and the recreational users  $u(s_i,x_j) = F(x_j) + G(s_i,x_j)$, depending on the water level $s_i$ and the amount of irrigation $x_i$ we get an auxiliary matrix

\begin{dmath}
aux &= \left[
 \begin{array}{ccccc} 
u(s_1, x_1)& \emptyset & \emptyset & ... & \emptyset \\ u(s_2, x_1)& u(s_2, x_2)& \emptyset & ... & \emptyset \\                                               u(s_3, x_1)& ... & u(s_3, x_3)& ... & \emptyset \\                                              ... & ... & ... & ... & \emptyset \\                                              u(s_n, x_1)& ... & ... & ... & u(s_n, x_n)\\                                               \end{array}\right]+\beta*\left[ \begin{array}{ccccc}
V'(s_1') & \emptyset & \emptyset & ... & \emptyset \\                                               V'(s_1') & V'(s_2') & \emptyset & ... & \emptyset \\                                               V'(s_1') & ... & V'(s_3') & ... & \emptyset \\                                              ... & ... & ... & ... & \emptyset \\                                              V'(s_1') & ... & ... & ... & V'(s_n') \\                                               \end{array}\right]
\end{dmath}

where the upper right triangle of this matrix will be empty, as the amount of water, that can be taken from the reservoir must be smaller or equal to the amount of water in the reservoir.

@@

\begin{equation}
V(s) = \max_{x}(u(s,x) + \beta*V'(s')))
\end{equation}

which equals to finding the column $aux$ that maximizes the expected value over all states (TODO: kann man das so sagen?). We repeat this step over and over again until the our value function $V$ converges (technically: the norm of the difference between $V'$ and $V$ falls below a certain tolerance value).

Finally, we want to compute the value for the steady state.
